---
title: "Seattle HIV Vaccine PrEP Model"
author: "Blythe Adamson"
date: "10/26/2018"
output: 
  html_document:
    theme: yeti
    highlight: tango
    toc: true
    toc_float: true
---

__Date generated:__ `r format(Sys.time(), '%d %B %Y')`

__Dynamic output used:__ May 30, 2018


***
# Overvie
***

### Purpose

This code analyzes output from a dynamic transmission model of HIV, add costs and quality of life adjustment, and performs sensitivity analyses. 

```{r setup, include=FALSE}

knitr::opts_chunk$set(fig.width = 10,
                      fig.height = 8,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

options(scipen = 999)

library(ggplot2)
library(reshape2)
library(dplyr)
library(RColorBrewer)
library(scales)
library(stringr)
library(kableExtra)

```

``` {r load data, cache = TRUE}

source('functions.R')

source('params.R')

source('costs.R')

source("costs.R")
    # if determinitic (probs = FALSE), then this produces a vector of costs
    # if probabilistic, this produces a matrix for the number of simulations

# load results from dynamic transmission model simulations

# reference group
hiv_ref <- read.csv(file = "../dynamic/raw/05Jun19/seattle_baseN.csv")

# primary result: 50% effectie vaccine
hiv_pv <- read.csv(file = "../dynamic/raw/05Jun19/seattle_vac50N.csv")

# sensitivity analysis: 70% effective vaccine
hiv_pv30 <- read.csv(file = "../dynamic/raw/05Jun19/seattle_vac30N.csv")

# sensitivity analysis: 70% effective vaccine
hiv_pv70 <- read.csv(file = "../dynamic/raw/05Jun19/seattle_vac70N.csv")

# scenario: no condom replacement, uses 50% efficacy
hiv_pvNC <- read.csv(file = "../dynamic/raw/05Jun19/seattle_vacNC.csv")

```

```{r pop size}
# POP SIZE --------------------------------------------------------------------

# Add variable for population size and year to each row
hiv_ref <- PopSize(hiv_ref, trt = ref)

hiv_pv  <- PopSize(hiv_pv, trt = pv)

hiv_pv70  <- PopSize(hiv_pv70, trt = pv)

hiv_pv30  <- PopSize(hiv_pv30, trt = pv)

hiv_pvNC  <- PopSize(hiv_pvNC, trt = pv)

n.i <- mean(hiv_ref$popsize) # average population size in reference case across years

mean(hiv_pv$popsize) - mean(hiv_ref$popsize) # mean difference in pop size, extra person-time

# Create a table of the population size changing over time
Case <- c("Reference: PrEP alone", "Intervention: Vaccine + PrEP")

Today_2018 <- c(hiv_ref$popsize[1],hiv_pv$popsize[1])

Year_2025 <- c(hiv_ref$popsize[hiv_ref$Time == 2025],hiv_pv$popsize[hiv_ref$Time == 2025])

End_2045 <- c(hiv_ref$popsize[hiv_ref$Time == 2045],hiv_pv$popsize[hiv_ref$Time == 2045])

Growth <- c(hiv_ref$popsize[hiv_ref$Time == 2045] - 
              hiv_ref$popsize[1],hiv_pv$popsize[hiv_ref$Time == 2045] - hiv_pv$popsize[1])

Percent_Growth <- c((hiv_ref$popsize[hiv_ref$Time == 2045] - hiv_ref$popsize[1])/hiv_ref$popsize[1],
                    (hiv_pv$popsize[hiv_ref$Time == 2045] - hiv_pv$popsize[1])/hiv_pv$popsize[1])

table_popsize <- cbind.data.frame(Case, Today_2018, Year_2025, End_2045, Growth, Percent_Growth)

```

```{r time horizon}
# TIME HORIZON ---------------------------------------------------------------
horizon_withcalibration <- max(hiv_ref$Time) - min(hiv_ref$Time)

horizon_withcalibration

horizon <- max(hiv_ref$Time) - 2025

```

The model time horizon was `r horizon` years, ending in year `r max(hiv_ref$Time)`.
There were `r hiv_pv$popsize[nrow(hiv_ref)] - hiv_ref$popsize[nrow(hiv_ref)]` additional people
alive at the end of the time horizon. 


#### Table. Population sizes at the start and end of the time horizon
```{r}
table_popsize %>%
  kable() %>%
  kable_styling()
```



## Health States and Utility

#### Table. Health states stratified by prevention, HIV status, CD4 count, treatment, and viral suppression

```{r health states}

# Define the attributes of each health state

state_names <- colnames(hiv_ref[,3:60])

states_n <- length(state_names)

source("StateAttributes.R") 

states <- DefineStates(state_names)  # data frame with indicator variables

save(states, file = "states.rda")

aware_labels <- unique(na.omit(states$aware))

cd4_labels <- unique(na.omit(states$cd4))

care_labels <- unique(na.omit(states$care))

# UTILITY --------------------------------------------------------------
# add utility weights for each health state
states <- UtilityWeights(states)

states %>%
  kable() %>%
  kable_styling()
```



```{r utilization}
# Probability of using PrEP vs Vaccines
hiv_ref$pPrEP <- rowSums(hiv_ref[,c(5,6,9,10)])/rowSums(hiv_ref[,3:10])
hiv_ref$pVacc <- rowSums(hiv_ref[,7:8])/rowSums(hiv_ref[,c(3:10)])

hiv_pv$pPrEP <- rowSums(hiv_pv[,c(5,6,9,10)])/rowSums(hiv_ref[,c(3:10)])
hiv_pv$pVacc <- rowSums(hiv_pv[,7:8])/rowSums(hiv_ref[,c(3:10)])

hiv_pv30$pPrEP <- rowSums(hiv_pv30[,c(5,6,9,10)])/rowSums(hiv_ref[,c(3:10)])
hiv_pv30$pVacc <- rowSums(hiv_pv30[,7:8])/rowSums(hiv_ref[,c(3:10)])

hiv_pv70$pPrEP <- rowSums(hiv_pv70[,c(5,6,9,10)])/rowSums(hiv_ref[,c(3:10)])
hiv_pv70$pVacc <- rowSums(hiv_pv70[,7:8])/rowSums(hiv_ref[,c(3:10)])

hiv_pvNC$pPrEP <- rowSums(hiv_pvNC[,c(5,6,9,10)])/rowSums(hiv_ref[,c(3:10)])
hiv_pvNC$pVacc <- rowSums(hiv_pvNC[,7:8])/rowSums(hiv_ref[,c(3:10)])
```


```{r outcomes}
# HEALTH OUTCOMES ------------------------------------------------------------

# LIFE YEARS (LY)-------------------------------------------------------------
hiv_ref_ly <- GetPersonTime(hiv_ref)  # see functions.R
hiv_pv_ly <- GetPersonTime(hiv_pv)    # see functions.R
hiv_pv30_ly <- GetPersonTime(hiv_pv30)    # see functions.R
hiv_pv70_ly <- GetPersonTime(hiv_pv70)    # see functions.R
hiv_pvNC_ly <- GetPersonTime(hiv_pvNC)    # see functions.R

# QALYs discounted-------------------------------------------------------------
hiv_ref_qaly <- GetQALYs(hiv_ref_ly)
hiv_pv_qaly  <- GetQALYs(hiv_pv_ly)
hiv_pv30_qaly  <- GetQALYs(hiv_pv30_ly)
hiv_pv70_qaly  <- GetQALYs(hiv_pv70_ly)
hiv_pvNC_qaly  <- GetQALYs(hiv_pvNC_ly)

# Collapse life years into annual and by care cascade categories
hiv_ref <- CareCascadeSummary(hiv_ref) # sums S and I compartments
hiv_pv <- CareCascadeSummary(hiv_pv)
hiv_pv30 <- CareCascadeSummary(hiv_pv30)
hiv_pv70 <- CareCascadeSummary(hiv_pv70)
hiv_pvNC <- CareCascadeSummary(hiv_pvNC)

# COSTS ------------------------------------------------------------

cPrep <- (1049.91*3) # FSS pricing of $1,049.91 from CCST in March 2018
  # cost for 3-months supply  
  
cVaccine <- (1623.72*(3/5)*1.3) #upfront one-time cost for 5-dose series of vaccine  
  # benchmarking reference: FSS price of Guardasil 9 is 1623.72 per dose
  # excludes delivery costs and labs

C <- GetCostParams(prob = FALSE, nsims = 1, cVaccine = cVaccine, cPrep = cPrep, societal = FALSE)

              # examine health state costs to check for errors
              #C 
              #dim(C)
              #length(C)
              #dim(hiv_ref_ly)

hiv_ref_cost <- CostingCalc(hiv_ref_ly, C)

hiv_pv_cost <- CostingCalc(hiv_pv_ly, C)

hiv_pv30_cost <- CostingCalc(hiv_pv30_ly, C)

hiv_pv70_cost <- CostingCalc(hiv_pv70_ly, C)

hiv_pvNC_cost <- CostingCalc(hiv_pvNC_ly, C)

```


# Results

```{r results tables}
# ------------------------------------------------------------------------------
# RESULTS 
# ------------------------------------------------------------------------------

#' Create results table for manuscript with HIV infections, PrEP use, life years, 
#'    QALYs, total cost, prep cost, vaccine cost, and ICER 

table_results <- GetResultsTable(hiv_ref, hiv_ref_ly, hiv_ref_qaly, hiv_ref_cost, 
                                 hiv_pv, hiv_pv_ly, hiv_pv_qaly, hiv_pv_cost)

table_results30 <- GetResultsTable(hiv_ref, hiv_ref_ly, hiv_ref_qaly, hiv_ref_cost, 
                                 hiv_pv30, hiv_pv30_ly, hiv_pv30_qaly, hiv_pv30_cost)

table_results70 <- GetResultsTable(hiv_ref, hiv_ref_ly, hiv_ref_qaly, hiv_ref_cost, 
                                 hiv_pv70, hiv_pv70_ly, hiv_pv70_qaly, hiv_pv70_cost)

table_resultsNC <- GetResultsTable(hiv_ref, hiv_ref_ly, hiv_ref_qaly, hiv_ref_cost, 
                                 hiv_pvNC, hiv_pvNC_ly, hiv_pvNC_qaly, hiv_pvNC_cost)

# Quantity of vaccine delivered during time horizon
series_delivered <- sum(hiv_pv$nVacc) # cumulative number of HIV vaccine series delivered

doses_delivered <- sum(hiv_pv$nVacc)*5 # cumulative doses delivered


hiv_prevalence_2018 <- hiv_ref %>%
  mutate(prevalence = sum_Infected/(sum_Susceptible + sum_Infected)) %>% 
  filter(Time == 2018) 
```


```{r Figure 5}
# new figures added Jan 10, 2019

ce_plane <- hiv_ref_cost %>% 
  mutate(cost_ref = totalCost) %>%   
  select(Time, cost_ref) %>% 
  left_join(hiv_ref_qaly %>% 
              mutate(qaly_ref = totalQALY) %>% 
              select(Time, qaly_ref), by = "Time") %>% 
  left_join(hiv_pv_cost %>% 
              mutate(cost_pv = totalCost) %>% 
              select(Time, cost_pv), by = "Time") %>% 
  left_join(hiv_pv_qaly %>% 
              mutate(qaly_pv = totalQALY) %>% 
              select(Time, qaly_pv), by = "Time") %>% 
  left_join(hiv_pv30_cost %>% 
              mutate(cost_pv30 = totalCost) %>% 
              select(Time, cost_pv30), by = "Time") %>% 
  left_join(hiv_pv30_qaly %>% 
              mutate(qaly_pv30 = totalQALY) %>% 
              select(Time, qaly_pv30), by = "Time") %>% 
    left_join(hiv_pv70_cost %>% 
              mutate(cost_pv70 = totalCost) %>% 
              select(Time, cost_pv70), by = "Time") %>% 
  left_join(hiv_pv70_qaly %>% 
              mutate(qaly_pv70 = totalQALY) %>% 
              select(Time, qaly_pv70), by = "Time") %>% 
  mutate(qaly_inc = qaly_pv - qaly_ref,
         cost_inc = cost_pv - cost_ref,
         qaly_30_inc = qaly_pv30 - qaly_ref,
         cost_30_inc = cost_pv30 - cost_ref,
         qaly_70_inc = qaly_pv70 - qaly_ref,
         cost_70_inc = cost_pv70 - cost_ref,
         #need the cumulative sum of costs and qalys
         qaly_inc_cum = cumsum(qaly_inc),
         cost_inc_cum = cumsum(cost_inc),
         qaly_30_inc_cum = cumsum(qaly_30_inc),
         cost_30_inc_cum = cumsum(cost_30_inc),
         qaly_70_inc_cum = cumsum(qaly_70_inc),
         cost_70_inc_cum = cumsum(cost_70_inc),
         # per person cumulative sum of costs and qalys
         qaly_inc_cum_pp = qaly_inc_cum/n.i,
         cost_inc_cum_pp = cost_inc_cum/n.i,
         icer = cost_inc_cum/qaly_inc_cum,
         lower = 50000*qaly_70_inc_cum,
         upper = 150000*qaly_70_inc_cum,
         pickyear = ifelse(Time == 2025 | Time == 2030 |Time == 2035 | Time == 2040 | 
                             Time == 2045, "True", "False"),
         year = as.character(floor(Time)))

ce_plane_brief <- ce_plane %>%
  filter(pickyear == "True")


ggplot(ce_plane_brief, aes(qaly_inc_cum, cost_inc_cum)) +
  geom_ribbon(aes(x = qaly_70_inc_cum, ymin = lower, ymax = upper, 
                  fill = "threshold"), alpha = .3) +
  geom_line(aes(qaly_70_inc_cum, lower), color = "grey") +
  geom_line(aes(qaly_70_inc_cum, upper), color = "grey") +
  
  geom_smooth(aes(color = "pv"), se = FALSE, size = 3) +
  geom_smooth(aes(qaly_70_inc_cum, cost_70_inc_cum, color = "pv70"), se = FALSE, 
              linetype = "dashed", size = 2) +  
  geom_smooth(aes(qaly_30_inc_cum, cost_30_inc_cum, color = "pv30"), se = FALSE, 
              linetype = "dashed", size = 2) +
  
  geom_point(aes(label = year), color = "dark red", size = 6) + 
  geom_point(aes(qaly_70_inc_cum, cost_70_inc_cum), color = "#00CCCC", size = 6) +
  geom_point(aes(qaly_30_inc_cum, cost_30_inc_cum), color = "orange", size = 6) +
  geom_text(data = ce_plane_brief %>% 
               filter(Time > 2034), aes(label=year),
            hjust = 0.7, vjust = -1.5, size = 6) +
  geom_text(data = ce_plane_brief %>% 
               filter(Time > 2034), aes(qaly_70_inc_cum, cost_70_inc_cum, label = year), 
            hjust = 0.5, vjust = -1.5, size = 6) +
  geom_text(data = ce_plane_brief %>% 
               filter(Time > 2034), aes(qaly_30_inc_cum, cost_30_inc_cum, label = year), 
            hjust = 1.5, vjust = 0, size = 6) +
  geom_text(data = ce_plane_brief %>% 
               filter(Time < 2034), aes(qaly_70_inc_cum, cost_70_inc_cum, label = year), 
            hjust = -.5, vjust = .5, size = 6) +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_continuous(labels = scales::comma) +
  labs(x = "QALYs Gained",
       y = "Cost to Implement HIV Vaccines with PrEP ($)") +
  scale_color_manual(values = c("pv30" = 'orange', "pv" = 'darkred', "pv70" = "#00CCCC"), 
                     labels = c("50%", "30%", "70%"), name = "HIV Vaccine Efficacy") +
  scale_fill_manual(values = c("threshold" = "grey70"), labels = c("1-3 x GDP per capita"), 
                    name = "Threshold for Cost-Effectiveness") +
  theme_bw() +  
  theme(text = element_text(size = 20), legend.position="bottom")

```



## Main Results: 50% efficacy

The model projects `r comma(series_delivered, 0)` vaccine HIV series delivered from 2025-2045, with a 
total of `r comma(doses_delivered, 0)` doses delivered. 

```{r}
table_results %>%
  kable() %>%
  kable_styling()
```

# Sensitivity Analysis 

## Results for HIV vaccine with 70% efficacy

```{r}
table_results70 %>%
  kable() %>%
  kable_styling()
```

## Results for HIV vaccine with 30% efficacy

```{r}
table_results30 %>%
  kable() %>%
  kable_styling()
```

## Scenario with no condom replacement

```{r}
table_resultsNC %>%
  kable() %>%
  kable_styling()
```



```{r sensitivity analysis, eval = FALSE}

# SENSITIVITY ANALYSIS ---------------------------------------------------------
cVaccine_base <- cVaccine
cPrep_base <- cPrep

cPrep <- cPrep_base*0.2
C <- GetCostParams(prob = FALSE, nsims = 1, cVaccine = cVaccine, cPrep = cPrep, societal = FALSE)
hiv_ref_cost <- CostingCalc(hiv_ref_ly, C)
hiv_pv_cost <- CostingCalc(hiv_pv_ly, C)
table_sa <- GetResultsTable(hiv_ref, hiv_ref_ly, hiv_ref_qaly, hiv_ref_cost, 
                                 hiv_pv, hiv_pv_ly, hiv_pv_qaly, hiv_pv_cost)
table_sa 

# one way sensitivity analysis
cVaccine <- cVaccine_base*0.575
C <- GetCostParams(prob = FALSE, nsims = 1, cVaccine = cVaccine, cPrep = cPrep, societal = FALSE)
hiv_ref_cost <- CostingCalc(hiv_ref_ly, C)
hiv_pv_cost <- CostingCalc(hiv_pv_ly, C)
table_sa <- GetResultsTable(hiv_ref, hiv_ref_ly, hiv_ref_qaly, hiv_ref_cost, 
                            hiv_pv, hiv_pv_ly, hiv_pv_qaly, hiv_pv_cost)
table_sa 

#HEAT MAP
v <- c(seq(from = 0, to = 3000, by = 120)) # cost of one vaccine dose

p <- c(seq(from = 0, to = 7500, by = 300)) # monthly cost of prep

sa <- matrix(c(rep(NA, length(v)*length(p))), nrow = length(v), ncol=length(p))

row.names(sa) <- v

colnames(sa) <- p

sa[1,] <- p

sa[,1] <- v

sa

for(x in 2:c(length(v))){
  for(y in 2:c(length(p))){
    cVaccine <- sa[x,1]
    cPrep <- sa[1,y]
    C <- GetCostParams(prob = FALSE, nsims = 1, cVaccine = cVaccine, cPrep = cPrep, societal = FALSE)
    hiv_ref_cost <- CostingCalc(hiv_ref_ly, C)
    hiv_pv_cost <- CostingCalc(hiv_pv_ly, C)
    #ICER sensitivity analysis
    sa[x,y] <- round((sum(hiv_pv_cost$totalCost[hiv_pv_cost$Time >= 2025]) 
                      - sum(hiv_ref_cost$totalCost[hiv_ref_cost$Time >= 2025]))
                     / (sum(hiv_pv_qaly$totalQALY[hiv_pv$Time>=2025]) 
                        - sum(hiv_ref_qaly$totalQALY[hiv_ref$Time>=2025]))) # ICER row
  }
}

#Reset to base case
cPrep <- (1049.91*3) # FSS pricing of $1,049.91 from CCST in March 2018

cVaccine <- (1623.72*(3/5)*1.3) #upfront one-time cost for 5-dose series of vaccine

C <- GetCostParams(prob = FALSE, nsims = 1, cVaccine = cVaccine, cPrep = cPrep, societal = FALSE)

hiv_ref_cost <- CostingCalc(hiv_ref_ly, C)

hiv_pv_cost <- CostingCalc(hiv_pv_ly, C)

head(sa)

dim(sa)

sa1 <- sa[2:dim(sa)[1],2:dim(sa)[2]]

head(sa1)

sa1 <- data.matrix(sa1)

save(sa1, file = "heatmap2.rda") # save the sensitivity analysis data

sa2 <- data.frame(sa1)

head(sa2)

mean(sa2$X600)

sa3 <- melt(data.frame(sa), id.vars = "X0")

sa3$cp <- as.numeric(gsub("X","",sa3$variable))/3 # 30-day supply of PrEP drugs

sa3$cv <- sa3$X0*5 # cost of compelete 5-dose vaccine series

sa3 <- sa3[sa3$cv != 0,]

head(sa3)
```

```{r threshold analysis, eval = FALSE}
# Threshold for cost-effectiveness

v5 <- v*5
v5 <- v5[-1]
wtp <- rep(150000, length(v5))
prepmax <- rep(NA,length(v5))
threshold <- data.frame(v5, wtp, prepmax)
threshold50 <- threshold
head(threshold)
for(i in v5){
  threshold$prepmax[v5 == i] <- min(sa3$cp[sa3$cv==i & sa3$value < 150000])
}
for(i in v5){
  threshold50$prepmax[v5 == i] <- min(sa3$cp[sa3$cv==i & sa3$value < 50000])
}
threshold <- threshold[threshold$prepmax != "Inf",]
threshold50 <- threshold50[threshold50$prepmax != "Inf",]
threshold50$prepmax[threshold50$prepmax == "Inf"] <- NA

fig_heatmap <- ggplot(sa3, aes(cv,cp)) +
  geom_tile(aes(fill = value), color = "white") + 
  scale_fill_gradient(low = "darkblue", high = "red", labels = comma) +
  geom_smooth(data = threshold, aes(x = v5, y = prepmax), se = FALSE, color = "gold") +
  geom_smooth(data = threshold50, aes(x = v5, y = prepmax), se = FALSE, 
              color = "gold", linetype = "dotted") +
  ylab("PrEP Cost (30-day supply)") +
  xlab("Vaccine Cost (per series)") +
  geom_point(aes(x=cVaccine*5, y=cPrep/3), colour="gold", size = 3) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 12),
        plot.title = element_text(size=16),
        axis.title=element_text(size=14,face="bold"),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(labels = scales::dollar) +
  scale_x_continuous(labels = scales::dollar) +
  labs(fill = "ICER ($/QALY)") +
  theme_bw() 
  
# "steelblue" color also looks nice

fig_heatmap

# ------------------------------------------------------------------------------
# ------------------------------------------------------------------------------
### TO DO LIST: compare distribution of cd4 counts to the NYC distributions
### show how much closer this gets us to eradication or 90:90:90 goals
```

